Implementing Distributed Systems Directly with I/O Automata in Ioa++

### 2011-09-22 papers/paper1/evaluation.tex
Revision to evaluation.

An automaton $R$ contains an input, output, and internal action.
The output and internal action execute a routine whose complexity is proportional to the parameter $N$.
The automaton executes a fixed number of local actions which was set to 1,000 for all experiments.
The automaton schedules the internal action with probability $\sqrt{\rho}$ and schedules the output action otherwise.
The system $S$ to be executed consists of two $R$ automata composed so the output action of one is composed with the input action of another.
If we divide the execution into rounds where each automaton executes a single action, the probability that both $R$ automata execute an internal action is $\sqrt{\rho} \times \sqrt{\rho} = \rho$.
Thus, we can use the parameter $\rho$ to vary the independence of the automata.

@@

@@

Figure~\ref{speed_up} shows the speed-up when system $S$ was executed with one and two threads.
The parameter $\rho$ was varied from $0.0$ to $1.0$ in increments of $0.1$.
The parameter $N$ was varied from $1$ to $1000000$ in factors of $10$.
When $\rho = 0$, every action is a bound output action and therefore depends on both $R$ automata.
Consequently, every action must be serialized yielding a maximum speed-up of 1.
When $\rho = 1$, every action is an internal action and can be executed concurrently with a corresponding maximum speed-up of 2.

@@

Figures~\ref{overhead1} and \ref{overhead2} show the per thread average of the fraction of time devoted to framework code and synchronization calls for the single and multi-threaded executions.
Since the number of actions to be executed is constant, the single threaded execution shows a relatively consistent fraction for each value of $N$.
The fraction decreases as $\rho$ goes to 1 as the system transitions from acquiring two locks to one lock.
The multi-threaded execution shows that the fraction of time devoted to system overhead decreases as $\rho$ goes to 1 and is more pronounced when $N$ is large.
For example, when $N=1000000$ one thread spends half of its time waiting on the other thread when $\rho = 0$ and spends virtually no time on synchronization when $\rho = 1$.

This experiment indicates that concurrent execution is possible with ioa++ but a speed-up will only be achieved if (1) enough independent actions are enabled and (2) the duration of the independent actions overcome the overhead of ioa++.
The overhead of ioa++ can be divided into three components corresponding to the time it takes to dispatch an action, the time required to synchronize using the pthreads library, and the time required to add an action to the scheduler via the \verb+ioa::schedule+ call.
From the results, we calculated the average overhead per action of the ioa++ framework when dispatching an action and found it to be 4211.5ns with a standard deviation of 765ns for the single threaded experiments and 4858ns with a standard deviation of 1099ns for the multi-threaded experiments.
Similary, we calculated the scheduling overhead and found it to be 227.5ns with a standard deviation of 62ns for the single threaded experiments and 27.9ns with a standard deviation of 79ns for the multi-threaded experiments.
Both of these can be improved with better scheduler design.
The synchronization time depends on the interactions among the system being executed, the scheduler, and the locking scheme used by the framework.
We plan to address scheduling and synchronization in future work.

>>>>

An automaton $R$ contains an input, output, and internal action.
The output and internal action execute a routine whose complexity is proportional to the parameter $N$.
The automaton executes a fixed number of local actions.
The automaton schedules the internal action with probability $\sqrt{\rho}$ and schedules the output action otherwise.
The system $S$ to be executed consists of two $R$ automata composed so the output action of one is composed with the input action of the other.
If we divide the execution into rounds where each automaton executes a single action, the probability that both $R$ automata execute an internal action is $\sqrt{\rho} \times \sqrt{\rho} = \rho$.
Thus, we can use the parameter $\rho$ to vary the independence of the automata.

@@

A trial consists of a choice for $\rho$, $N$, and the number of actions to be executed.
For the experiments, the parameter $\rho$ was varied from $0.0$ to $1.0$ in increments of $0.1$ and the parameter $N$ was varied from $1$ to $1000000$ by factors of $10$.
The number of actions executed by each automaton was fixed at 1000.
Each trial was repeated 30 times.
All calculations were performed assuming a normal distribution at a confidence level of 95\%.

The driver for a single trial can be found in examples/random.cpp of the ioa++ package.
The trials were performed on a MacBook Pro running OS X 10.5.8 with an 2.4 GHz Intel Core 2 Duo processor with 4GB of memory.
The code was compiled and linked with i686-apple-darwin9-g++-4.0.1 build 5493 with -O2 optimization and the -DPROFILE flag to enable profiling in ioa++.

@@

Figure~\ref{speed_up} shows the speed-up when system $S$ was executed with one and two threads.
When $\rho = 0$, every action is a bound output action and therefore depends on both $R$ automata.
Consequently, every action must be serialized yielding a maximum speed-up of 1.
When $\rho = 1$, every action is an internal action and can be executed concurrently with a corresponding maximum speed-up of 2.

@@

Figures~\ref{overhead1} and \ref{overhead2} show the per thread average of the fraction of time devoted to framework code and synchronization calls for the single and multi-threaded executions.
(Note that the confidence intervals are present but small.)
Since the number of actions to be executed is constant, the single threaded execution shows a relatively consistent fraction for each value of $N$.
The fraction decreases as $\rho$ goes to 1 as the system transitions from acquiring two locks to one lock.
The multi-threaded execution shows that the fraction of time devoted to system overhead decreases as $\rho$ goes to 1 and is more pronounced when $N$ is large.
For example, when $N=1000000$ one thread spends half of its time waiting on the other thread when $\rho = 0$ and spends virtually no time on synchronization when $\rho = 1$.

This experiment indicates that concurrent execution is possible with ioa++ but a speed-up will only be achieved if (1) enough independent actions are enabled and (2) the duration of the independent actions overcome the overhead of ioa++.
The overhead of ioa++ can be divided into three components corresponding to the time it takes to dispatch an action, the time required to synchronize using the pthreads library, and the time required to add an action to the scheduler via the ioa::schedule call.
From the results, we calculated the average overhead per action of the ioa++ framework when dispatching an action and found it to be 4211ns $\pm$ 31ns for the single threaded experiments and 4858ns $\pm$ 32ns for the multi-threaded experiments.
Similary, we calculated the scheduling overhead and found it to be 227ns $\pm$ 2.5ns for the single threaded experiments and 279ns $\pm$ 2.3ns for the multi-threaded experiments.
Both of these can be improved with better scheduler design.
The synchronization time depends on the interactions among the system being executed, the scheduler, and the locking scheme used by the framework.
We plan to address scheduling and synchronization in future work.
